\sffamily An EEG-based visual Brain-Computer Interface for gaze-free communication

### 2024-09-30 chapters/gaze_independence.tex
New patient table

@@

\todo{check funding proposal}
\todo{check provisional doctoral plan}
\todo{check WIP seminar}

@@

\todo{mention, free, split, frenzel}
\todo{Overt, covert, split, free, illustration from poster}
\todo{clear definitions, table from paper}

>>>>

\todo{
14] Fazzi E, Signorini SG, Piana RL, et al. Neuro-ophthalmo-
logical disorders in cerebral palsy: ophthalmological, oculo-
motor, and visual aspects. Dev Med Child Neurol. 2012;54:
730–736.
[15] Graber M, Challe G, Alexandre MF, et al. Evaluation of the
visual function of patients with locked-in syndrome: report
of 13 cases. J Fr Ophtalmol. 2016;39:437–440.
[16] Moss HE, McCluskey L, Elman L, et al. Cross-sectional evalu-
ation of clinical neuro-ophthalmic abnormalities in an
amyotrophic lateral sclerosis population. J Neurol Sci. 2012;
314:97–101.
}

@@

\todo{
  Complete with refs from

Fast detection of covert visuospatial
attention using hybrid N2pc and SSVEP
features
}

@@

\todo{mention, free, split, frenzel}
\todo{clear definitions, table from paper}

### 2024-09-30 chapters/patients.tex
New patient table

Communication BCIs have a target population consisting of patients in various
stages of paralysis or \ac{lis}, often suffering from oculomotor impairments.
This reduces their performance operating visual oddball \ac{bci}s (see
section\todo{ref section} for an overview), since they
are presumed unable to comfortably redirect their gaze at the desired target,
i.e., perform overt \ac{vsa}.\todo{cite some visual gaze-independent bci}
They might more naturally operate in covert \ac{vsa}, where the gaze and
\ac{vsa} do not coincide.
Several studies show that performance drops when not fixating the intended
target~\cite{Brunner2010, Frenzel2011, Treder2010, RonAngevin2019,
VanDenKerchove2024}, necessitating gaze-independent solutions.

@@

Our previous study, presented in chapter \todo{ref chapter}  partially tried to account for this
and showed gaze-independent performance can be improved
improved in healthy subjects using a suited decoding strategy.
Yet, there is a strong need for verification of results obtained in healthy
samples.
It is a striking constatation that studies reporting on
gaze-independent visual \ac{bci} in patients that are actually eye-motor
impaired are very few.
in patients, cite different characteristics, differences
in setting, environment and equipment,
\todo{List conclusions of Peters2020 (severely eye motor impaired als patients
able to use a hex-o-spell)}

@@

@@

>>>>

Communication BCIs have a target population consisting of patients in various
stages of paralysis or \ac{lis},
severe motor impairment
\todo{cite and discuss Usability and Workload of Access
Technology for People With Severe Motor
Impairment: A Comparison of Brain-
Computer Interfacing and Eye Trackin}

often suffering from oculomotor impairments.
can introduce problems in bci ~\cite{Brain-computer interface (BCI) evaluation
in people with amyotrophic lateral sclerosis}, discuss als example

\todo{. As with eye tracking SGDs, visual
and oculomotor impairments may limit PALS’s performance
with visual BCI systems (Fried-Oken et al., 2020). Although
}

\cite{FriedOken2020}

This reduces their performance operating visual oddball \ac{bci}s (see
section\todo{ref section} for an overview), since they
are presumed unable to comfortably redirect their gaze at the desired target,
i.e., perform overt \ac{vsa}.\todo{cite some visual gaze-independent bci}
They might more naturally operate in covert \ac{vsa}, where the gaze and
\ac{vsa} do not coincide.
Several studies show that performance drops when not fixating the intended
target~\cite{Brunner2010, Frenzel2011, Treder2010, RonAngevin2019,
VanDenKerchove2024}, necessitating gaze-independent solutions.

@@

Our previous study, presented in chapter \todo{ref chapter}  partially tried to account for this
and showed gaze-independent performance can be improved
improved in healthy subjects using a suited decoding strategy.
Yet, there is a strong need for verification of results obtained in healthy
samples.
It is a striking constatation that studies reporting on
gaze-independent visual \ac{bci} in patients that are actually eye-motor
impaired are very few.
in patients, cite different characteristics, differences
in setting, environment and equipment,
\todo{List conclusions of Peters2020 (severely eye motor impaired als patients
able to use a hex-o-spell)}
\todo{discuss Lesenfants2014}
\todo{discuss Marchetti2013}

@@

Self-reported according to cite.
Additionally, the participants were asked if they experienced any tremors like
nystagmus or other (larger) involuntary eye movements.

@@

The Coma Recovery Scale
Revised (CRS-R) [73] provides standardized procedures for admin-
istration and response criteria to more uniformly describe the vis-
ual fixation or pursuit skills of a user with SSPI. The Northeastern
State University College of Optometry (NSUCO) Oculomotor Test
[74] provides specific procedural techniques as well as scoring cri-
teria, which allows for uniform description of the accuracy and
ability of pursuit and saccade skills for adults and children.
